Algebraic filtering of surfaces from medical images with Julia

### 2019-11-10 src/paper.tex
paper revised

\paragraph{\emph{Chain coordinates }\label{sec:bbbb}}
We are going to treat each image block independently from each other. Hence we map each image block $\B(i,j,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n$, using coordinate vectors $c_{h,k}\in \B^{n\times n} := \{0,1\}^{n\times n}$, where the basis element $c = c_{h,k} \in C_2$ is mapped via Cartesian-to-linear map to the binary vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \B^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

@@

\paragraph{\emph{Boundary operator }\label{sec:bbbb}}For a fixed block size $n$, the boundary operator $\partial_d : C_d\to C_{d-1}$, with $d\in\{2,3\}$, will be constructed once and for all using the algorithm given in~\cite{}, and inlined in the generated boundary extraction code.

@@

\paragraph{\emph{Sparsity and size of boundary matrix }\label{sec:bbbb}}

@@

\paragraph{Embedding}
A final computational step is needed, in order to embed the 2-chains in Euclidean space $\E^3$ and to assemble the whole resulting surface. In particular, we need to compute the \emph{embedding function} $\mu : C_0 \to \E^3$, where $C_0$ is the space of 0-chains, one-to-one corresponding to the vertices of the extracted surface. The simplest solution is to associate  four 0-cells to each 2-cell of the extracted surface, i.e.~to each non-zero entry in every column of $\T{\B}$.  The $\mu$ function  can be computed by identifying, via  element position in the column, a triple of integer values $0\leq x\leq u$, $0\leq y\leq v$, and $0\leq z\leq w$ for each vertex of the 2-cell.  The mapping can be implemented using a dictionary, that will store the inverse coordinate transformation used at the beginning, i.e.~the one from linear to Cartesian coords, in order of not duplicating the output vertices.

\paragraph{Surface assembling}

@@

\paragraph{\emph{Workflow setup}}The functions in this preliminary step include:
\begin{enumerate}

@@

\paragraph{\emph{Job enqueuing}}Communication and data synchronization may be managed through \emph{Channels}, which are the FIFO conduits that may provide producer/consumer communication. Overall execution time can be improved if other tasks can be run while a task is being executed, or while waiting for an external service/function to complete. The single work items of this stage follow:
\begin{enumerate}

@@

\paragraph{\emph{3-Chain encoding}}
The interesting part of the \emph{Image} $\mathcal{I}$ is called \emph{Segment} $\mathcal{S}$. The goal of the whole \emph{workflow} is to extract a \emph{boundary model} of $\mathcal{S}$ from $\mathcal{I}$. The portion of $\mathcal{S}$ inside $\mathcal{B}$, will be denoted as $\mathcal{S}(B)$.
Each block $\mathcal{B}$ of the 3D image must by converted into the \emph{coordinate representation} of a vector $\nu\in C_3$ in the linear space  of 3-chains.

@@

\paragraph{\emph{SpMM Multiplication}}
According to the current literature~\cite{} it is more convenient to execute SpMV (sparse matrix-vector) multiplications than SpMSpV (sparse matrix-sparse vector) multiplications. Since we have 256 such jobs (one multiplication per block) to perform in the standard setting of the algorithm\footnote{
Size of block $64^3$; size of image $512^2\times 256$.
}, or more in case of either smaller blocks or image greater than the standard one, this stage must be evidently parallelized and carefully tuned, possibly by using the GPU, if available.
\begin{enumerate}

@@

\paragraph{\emph{2-Chain decoding}}
Each multiplication of $[\partial_B] : C_3 \to C_2$, times a 3-chain $\nu\in C_3$, produces a 2-chain  $\sigma\in C_2$, i.e.~the \emph{coordinate representation} of the \emph{boundary vector} $\sigma\in C_2$.  The inverse of the coding algorithm is executed in the present stage.  This process can also be partially superimposed in time with the previous ones, depending on the size of the memory buffers used to feed the CPU cores or the GPUs and get their results. Some elementary steps follow:

@@

\paragraph{\emph{Assembling and artifact filtering}}
The results of the previous stages can be described as a \emph{collection of sets} of \emph{geometric quadrilaterals (quads)}, each one encoded as an array of quadruples of integer indices, pointing to the linear array of grid vertices associated to the image block $\mathcal{B}$.  In other words, \emph{all quads} of \textbf{each job} are now given in the \textbf{same} \emph{local coordinates}.  Besides to put each partial surface $\mathcal{S}(B) = (\texttt{V}_B, \texttt{FV}_\sigma)$ in the global coordinate system of the image, the present stage must eliminate the redundant boundary features possibly generated at the edges of the partial surface $\mathcal{S}(B)$ within each block $\mathcal{B}$ such that $B \cap \mathcal{I} \not= \emptyset$:

@@

\paragraph{\emph{Smoothing}}
The final smoothing of the generated surfaces cannot be performed block-wise, since this would introduce smoothing artifacts at the block boundaries. Anyway, the Taubin smoothing~\cite{} can be performed in parallel, since for each vertex in the final surface (except eventually the ones on the image $\mathcal{I}$ boundaries) it essentially consist in computing a new position as a proper average of its neighborhood vertices, i.e.~by applying a discrete Laplacian operator.  Some appropriate set workers may so be assigne the task of generating iteratively a new position for the vertices they take cure of. In particular, we have:
\begin{enumerate}

>>>>

\subsubsection{Chain coordinates }\label{sec:bbbb}
We are going to treat each image block independently from each other. Hence we map each image block $\B(i,j,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n$, using coordinate vectors $c_{h,k}\in \B^{n\times n} := \{0,1\}^{n\times n}$, where the basis element $c = c_{h,k} \in C_2$ is mapped via Cartesian-to-linear map to the binary vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \B^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

@@

\subsubsection{Boundary operator }\label{sec:bbbb}
For a fixed block size $n$, the boundary operator $\partial_d : C_d\to C_{d-1}$, with $d\in\{2,3\}$, will be constructed once and for all using the algorithm given in~\cite{}, and inlined in the generated boundary extraction code.

@@

\subsubsection{Sparsity and size of boundary matrix }\label{sec:bbbb}

@@

\subsubsection{Embedding}
A final computational step is needed, in order to embed the 2-chains in Euclidean space $\E^3$ and to assemble the whole resulting surface. In particular, we need to compute the \emph{embedding function} $\mu : C_0 \to \E^3$, where $C_0$ is the space of 0-chains, one-to-one corresponding to the vertices of the extracted surface. The simplest solution is to associate  four 0-cells to each 2-cell of the extracted surface, i.e.~to each non-zero entry in every column of $\T{\B}$.  The $\mu$ function  can be computed by identifying, via  element position in the column, a triple of integer values $0\leq x\leq u$, $0\leq y\leq v$, and $0\leq z\leq w$ for each vertex of the 2-cell.  The mapping can be implemented using a dictionary, that will store the inverse coordinate transformation used at the beginning, i.e.~the one from linear to Cartesian coords, in order of not duplicating the output vertices.

\subsubsection{Surface assembling}

@@

\subsubsection{Workflow setup}\label{}The functions in this preliminary step include:
\begin{enumerate}

@@

\subsubsection{Job enqueuing}\label{}Communication and data synchronization may be managed through \emph{Channels}, which are the FIFO conduits that may provide producer/consumer communication. Overall execution time can be improved if other tasks can be run while a task is being executed, or while waiting for an external service/function to complete. The single work items of this stage follow:
\begin{enumerate}

@@

\subsubsection{3-Chain encoding}\label{}
The interesting part of the \emph{Image} $\mathcal{I}$ is called \emph{Segment} $\mathcal{S}$. The goal of the whole \emph{workflow} is to extract a \emph{boundary model} of $\mathcal{S}$ from $\mathcal{I}$. The portion of $\mathcal{S}$ inside $\mathcal{B}$, will be denoted as $\mathcal{S}(B)$.
Each block $\mathcal{B}$ of the 3D image must by converted into the \emph{coordinate representation} of a vector $\nu\in C_3$ in the linear space  of 3-chains.

@@

\subsubsection{SpMM Multiplication}\label{}
According to the current literature~\cite{} it is more convenient to execute SpMV (sparse matrix-vector) multiplications than SpMSpV (sparse matrix-sparse vector) multiplications. Since we have 256 such jobs (one multiplication per block) to perform in the standard setting of the algorithm\footnote{
Size of block $64^3$; size of image $512^2\times 256$.
}, or more in case of either smaller blocks or image greater than the standard one, this stage must be evidently parallelized and carefully tuned, possibly by using the GPU, if available.
\begin{enumerate}

@@

\subsubsection{2-Chain decoding}\label{}
Each multiplication of $[\partial_B] : C_3 \to C_2$, times a 3-chain $\nu\in C_3$, produces a 2-chain  $\sigma\in C_2$, i.e.~the \emph{coordinate representation} of the \emph{boundary vector} $\sigma\in C_2$.  The inverse of the coding algorithm is executed in the present stage.  This process can also be partially superimposed in time with the previous ones, depending on the size of the memory buffers used to feed the CPU cores or the GPUs and get their results. Some elementary steps follow:

@@

\subsubsection{Assembling and artifact filtering}\label{}
The results of the previous stages can be described as a \emph{collection of sets} of \emph{geometric quadrilaterals (quads)}, each one encoded as an array of quadruples of integer indices, pointing to the linear array of grid vertices associated to the image block $\mathcal{B}$.  In other words, \emph{all quads} of \textbf{each job} are now given in the \textbf{same} \emph{local coordinates}.  Besides to put each partial surface $\mathcal{S}(B) = (\texttt{V}_B, \texttt{FV}_\sigma)$ in the global coordinate system of the image, the present stage must eliminate the redundant boundary features possibly generated at the edges of the partial surface $\mathcal{S}(B)$ within each block $\mathcal{B}$ such that $B \cap \mathcal{I} \not= \emptyset$:

@@

\subsubsection{Smoothing}\label{}
The final smoothing of the generated surfaces cannot be performed block-wise, since this would introduce smoothing artifacts at the block boundaries. Anyway, the Taubin smoothing~\cite{} can be performed in parallel, since for each vertex in the final surface (except eventually the ones on the image $\mathcal{I}$ boundaries) it essentially consist in computing a new position as a proper average of its neighborhood vertices, i.e.~by applying a discrete Laplacian operator.  Some appropriate set workers may so be assigne the task of generating iteratively a new position for the vertices they take cure of. In particular, we have:
\begin{enumerate}
